---
title: "Assignment1- Team7"
authors: " Maria Ahumada, Harsh Harsh, Linus Rupp, Melissa Talgo, Pranjal Srivastava"
date: "01/28/2021"
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: pdflatex
geometry: margin=3cm
fontsize: 11pt
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

```{r package installations, include=FALSE,warning=FALSE}
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)
install.packages('knitr')
install.packages('gsubfn')
install.packages('dplyr')
install.packages("readxl")
install.packages("tidyr")
install.packages('ggplot2')
install.packages(c("forecast", "TTR"), dependencies=TRUE)
```

```{r include=FALSE,warning=FALSE,tidy=TRUE}
library(knitr)
library(gsubfn)
library(dplyr)
library(readxl)
library(tidyr)
library(ggplot2)
library(forecast)
library(TTR)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

# Part 1
## Read the data & select 5 variables to perform decomposition
```{r part1, echo=TRUE, tidy=TRUE}
rm(list=ls(all=TRUE))
#Input data
data<-read.csv("data_week.csv", sep=",",dec=".",header=T) 	

#Preview of the data
head(data,5)

#variable #1
sales = (data$netsales_online)/1000000		# remove the millions so it becomes a small digit
head(sales)
#yy = ts(sales, frequency = 52,start = c(2015,1))
#plot.ts(yy)

#variable #2
snow_level <- data$avg_snow_level #numeric data type
head(snow_level)

#variable #3
#rainfall
avg_rainfall_level <- data$avg_rainfall_level
head(avg_rainfall_level)

#variable #4
#budget
budget <- data$online_budget/100000
head(budget)

#variable #5
#cloud_index
cloud_index <- data$avg_cloud_index
head(cloud_index)
```


# Part 2
## Represent data as a time-series object
```{r part2, echo=TRUE,tidy=TRUE}
##### Representing Data as Time Series Objects #####
# convert [x variable] into an R time series object; weekly means frequency is 52 and starting in year 2015, week 01
convert_to_ts <-  function(explanatory_variable){
  x <- ts(explanatory_variable,frequency = 52, start = c(2015,1))
  return(x)
}

#Apply function to all 5 variables chosen
sales_ts <- convert_to_ts(sales)
snow_level_ts <- convert_to_ts(snow_level)
avg_rainfall_level_ts <- convert_to_ts(avg_rainfall_level)
budget_ts <- convert_to_ts(budget)
cloud_index_ts <- convert_to_ts(cloud_index)

#Plot time series
plot.ts(sales_ts)
plot.ts(snow_level_ts)
plot(avg_rainfall_level_ts)
plot(budget_ts)
plot(cloud_index_ts)
```


# Part 3
## Time Series Decomposition into Signal+ Noise where Signal = Trend + Seasonality
```{r part3, echo=TRUE,tidy=TRUE}
decompose_ts_variable <- function(ts_variable){
  decomp =  decompose(ts_variable)
  trend = decomp$trend
  seasonal = decomp$seasonal
  seasonal_adj = ts_variable - seasonal
  random = decomp$random
  plot.ts(cbind(ts_variable,trend, seasonal,seasonal_adj, random))
}

#Apply function to all 5 variables we need to decompose
decompose_ts_variable(sales_ts)
decompose_ts_variable(snow_level_ts)
decompose_ts_variable(avg_rainfall_level_ts)
decompose_ts_variable(budget_ts)
decompose_ts_variable(cloud_index_ts)
```


# Part 4 #

## Holt-Winters ##

```{r part 4, echo=TRUE,tidy=TRUE}
calculate_holt_winters <- function(ts_variable, ts_var_name){
output1 = HoltWinters(ts_variable, beta=TRUE, gamma=TRUE) #both trend and seasonality ON
output2 = HoltWinters(ts_variable, beta=FALSE, gamma=FALSE) #both trend and seasonality OFF
output3 = HoltWinters(ts_variable, beta=TRUE, gamma=FALSE) #trend ON and seasonality OFF
output4 = HoltWinters(ts_variable, beta=FALSE, gamma=TRUE) #trend OFF and seasonality OFF


#### plotting the fitted method ###

par(mfrow=c(2,2))
plot(output1, main=paste("HW(W gamma and beta) for",ts_var_name, sep = " ", collapse = NULL))
plot(output2, main=paste("HW (W/O gamma and beta) for",ts_var_name, sep = " ", collapse = NULL))
plot(output3, main=paste("HW (W/O gamma and W beta) for",ts_var_name, sep = " ", collapse = NULL))
plot(output4, main=paste("HW (W gamma and W/O beta) for",ts_var_name, sep = " ", collapse = NULL))


#### plotting the forecasted method #####

autoplot(ts_variable) +
  autolayer(forecast(output1, h = 26), series="Holt-Winter's (With gamma and beta)", PI=FALSE) +
  autolayer(forecast(output2, h = 26), series="Holt-Winter's (Without gamma and beta)", PI=FALSE) +
  autolayer(forecast(output3, h = 26), series="Holt-Winter's (With beta and without gamma)", PI=FALSE) +
  autolayer(forecast(output4, h = 26), series="Holt-Winter's (Without beta and with gamma)", PI=FALSE) +
  ylab(ts_var_name) + xlab("Year") + 
  guides(colour=guide_legend(title="Forecast")) +
  ggtitle("Comparing different version of Holt's Method") + xlab("Year") 


}

calculate_holt_winters(sales_ts, "Sales_ts")
calculate_holt_winters(snow_level_ts, "snow_level_ts")
calculate_holt_winters(avg_rainfall_level_ts, "avg_rainfall_level_ts")
calculate_holt_winters(budget_ts, "budget_ts")
calculate_holt_winters(cloud_index_ts, "cloud_index_ts")
```






# Part 6
## Check residuals for the four combinations using Holts-Winter Filter
```{r part6, echo=TRUE,tidy=TRUE}

checking_residuals <- function(ts_variable, ts_var_name){
output1 = HoltWinters(ts_variable, beta=TRUE, gamma=TRUE) #both trend and seasonality ON
output2 = HoltWinters(ts_variable, beta=FALSE, gamma=FALSE) #both trend and seasonality OFF
output3 = HoltWinters(ts_variable, beta=TRUE, gamma=FALSE) #trend ON and seasonality OFF
output4 = HoltWinters(ts_variable, beta=FALSE, gamma=TRUE) #trend OFF and seasonality OFF

out1_res <- residuals(output1)
out2_res <- residuals(output2)
out3_res <- residuals(output3)
out4_res <- residuals(output4)

checkresiduals(out1_res)
checkresiduals(out2_res)
checkresiduals(out3_res)
checkresiduals(out4_res)

}

```



# Part 7
## Retain only one out of four models per variable decomposed -- give your justification
```{r part7, echo=TRUE}

checking_residuals(sales_ts, "sales_ts")
checking_residuals(snow_level_ts, "snow_level_ts")
checking_residuals(avg_rainfall_level_ts, "avg_rainfall_level_ts")
checking_residuals(budget_ts, "budget_ts")
checking_residuals(cloud_index_ts, "cloud_index_ts")

```

### Sales_ts ###

I believe for sales ts we can directly ignore Holt-Winter's method with either the gamma or beta missing since residuals are clearly not normal. For method with both beta and gamma true, I the correlation with lagged residuals is very high after the 6th lag, whereas in the ACF chart for Holt-winters without beta or gamma, the lagged residuals start dominating only after the 22nd term, and hence, we will choose Holt-Winter's method without beta and gamma to forecast sales_ts.
